---
title: "pca_shuhan"
author: "Shuhan Song"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
```

## 1. Get data
```{r}
ca_pb <- read_csv(here("data", "ca_pollution_burden.csv"))
ca_dem <- read_csv(here("data", "ca_census_demographics_2010.csv"))
```


## 2. Clean data and wrangling

```{r}
ca_pb_nopct <- ca_pb %>% 
  clean_names() %>% 
  select(-contains("perc")) %>% 
  select(-contains("pctl")) %>% 
  select(-latitude, -longitude)
```

```{r}
ca_dem_clean <- ca_dem %>% 
  clean_names()
```

## 3. Pollution burden indicators only

```{r}
ca_pb_subset <- ca_pb_nopct %>% 
  select(ozone:solid_waste, asthma:housing_burden)
```

Let's do PCA

```{r, eval=FALSE}
# gives error because there are missings
pb_pca <- prcomp(ca_pb_subset, scale = TRUE)
```

Aside: exploring missingness

```{r}
summary(ca_pb_subset) # 242 is relatively small number of missing compared with 8000+ observations

gg_miss_var(ca_pb_subset)

matrixplot(ca_pb_subset) # can be used to study correlation between variables, dark is higher numbers, red is missing numbers 

matrixplot(ca_pb_subset, sortby = "poverty") # sortby ordered poverty magnitude from low to high
```

Remove any observation with ANY NA values

```{r}
ca_pb_nona <- ca_pb_subset %>% 
  drop_na()

skim(ca_pb_nona)
```

Now, no NA values. otherwise, consider imputation, etc. to do pca

```{r}
my_ca_pca <- prcomp(ca_pb_nona, scale = TRUE)

my_ca_pca

summary(my_ca_pca)

biplot(my_ca_pca) # not a useful plot

# use autoplot!

my_biplot <- autoplot(my_ca_pca,
                      colour = NA, # point color
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label = TRUE,
                      loadings.label.repel = TRUE) +
  scale_y_continuous(limits = c(-0.02, 0.02))

my_biplot
```

## 4. PCA for pollution burden and demographics

join together ca_dem_clean with ca_pb_nopct

```{r}
ca_df <- ca_dem_clean %>% 
  inner_join(ca_pb_nopct, by = c("census_tract_number" = "census_tract"# can add other things to match separated by comma ",", the colume names' order matters here
                                 ))

ca_nona <- ca_df %>% drop_na()
```

Make a subset of ca_nona that only includes: 

- white_percent 
- elderly_65_percent
- pm2_5
- pesticides
- traffic
- asthma
- cardiovascular_disease
- poverty

```{r}
my_sub <- ca_nona %>% 
  select(white_percent, elderly_65_percent, pm2_5, pesticides, traffic, asthma, cardiovascular_disease, poverty)

```


```{r}
my_dem_pca <- prcomp(my_sub, scale = TRUE)

my_biplot_dem <- autoplot(my_dem_pca,
                      colour = NA, # point color
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label = TRUE,
                      loadings.label.repel = TRUE) +
  scale_y_continuous(limits = c(-0.05, 0.05)) +
  theme_bw()

my_biplot_dem
```

